Apparent preferences for cognitive effort fade when multiple forms of effort

### 2020-11-11 Cost_manuscript.Rmd
adding discussion points

  Understanding the sporadic attractiveness of effortful options remains elusive. In this study, we tested the hypothesis that such appeal could arise from being able to compare different means to achieve the same outcome. Using two foraging experiments that exposed individuals to a single or intermixed types of demands, we found that 1) preferences for cognitive effort are stable when individuals face a single form of demand, but fade over time when other costs are also experienced; and 2) the phenomenon can be explained by a model in which preferences depend on how participants construe the amount of time a task takes, which could hypothetically be recalibrated when all tasks are interleaved.

  The observed choice dynamics defied a number of explanatory candidates. For instance, differences in choice stability between experiments suggest that a demand's cost is not intrinsic, while differences between effort types rule out learned industriousness. Effort preferences due to boredom would not have shown the global acceptance decline and convergence in the within-subjects experiment, which also discards the possibility that all our participants were high on need for cognition (CITE). Even though fatigue could explain the decrease in effort acceptances across experiments, we would have expected tired participants to turn to wait trials over time in order to preserve earning rates. We also ruled out the sunk cost effects used to explain overharvesting (Wikenheiser2013; Sweiss2018), as participants expressed their choices as early as possible, resulting in occassional underharvesting. Even though the reciprocal relationship between effort and outcome can often explain preferences for effort (Inzlicht2018), the observed differences in acceptances between effort types and the eventual convergence of rates in the within-subjects study cannot be explained by reward alone, as demands were rewarded at the same rate. Instead, we propose that features of the cognitive tasks added outcome-independent value to the offer. Subjective time sensitivity has been shown to play a role in temporal discounting, improving on classic hyperbolic fits (Mckelachar....) *add that stability meant no need to model asymmetric learning across environmental richness from Neal*

- The optimality of the strategies produced by each cost depended on the richness of the block. This can help explain why the wait group earned the most despite having similar acceptance rates to the physical group, as the former was closer to optimality in the richest environment.
- Make sure to note that we controlled for time as a confound for effort, as more effortful tasks normally take longer
- What we think: subjective timing is shortened by the way we implemented cognitive effort. This can be recalibrated if many task types are experienced. Maybe the implementation allowed for counting. Alternatively, working memory engagement could produce timing issues (Buhi n Gershman). Also look at what Andrew Westbrook shared. Add this to a potential fatigue, although that would also mean that waiting would be accepted less over time in exp 1.
- Things that don't seem to be at play here, but could still play a general role in mental effort preferences: 1) A demand’s cost is not fully intrinsic (within vs between differences); 2) Effort preferences unlikely just due to learned industriousness or efficacy as amitais lab sees it (differences between effort types, and everyone met a good performance level per training); 3) avoiding boredom (global decline in exp 2, not a preference exchange). 4) Literally counting tasks, since mental effort was discretely presented but everything else was contiguous. If that were the case, dividing the handling time by 2 would work, but not sure if it would translated into exp 2. 5) fatigue: there's a similar gradual decrease in acceptances in both experiments, except for the wait condition in exp 1. Fatigue would apply equally to both, or alternatively it might not bring all down to the same level (I guess the point could be made, but it's not that people who complete more at the beginning decrease acceptances more, as one would expect from fatigued people). Also worth mentioning that even if this is due to need for cognition (i.e. Ceyda; Laura; Cacioppo et al 1996), the collapsing in the second experiment shows that preferences for cognitive effort cannot be just that.
- Deviations in the willingness to pursue effort can predict depression relapse (https://jamanetwork.com/journals/jamapsychiatry/article-abstract/2761562), and motivational factors affecting effort aversion can be heterogenous between depression and schizophrenia (https://www.researchgate.net/publication/337361164_Pathways_to_Motivational_Impairments_in_Psychopathology_Common_Versus_Unique_Elements_Across_Domains). Like, the aversion is due to focusing on different elements of the task.
- Talk about future experiments: if this is subjective time, changing the pacing of the tasks should change the acceptance rates for cognitive effort
- It's true that we didnt calibrate cog effort per subject. But instead we brought all subjects to the same level of performance, and we saw that as time progressed they became less likely to accept longer periods of effort. 
- Clinical: Deviations in the willingness to pursue effort can predict depression relapse (https://jamanetwork.com/journals/jamapsychiatry/article-abstract/2761562), and motivational factors affecting effort aversion can be heterogenous between depression and schizophrenia (https://www.researchgate.net/publication/337361164_Pathways_to_Motivational_Impairments_in_Psychopathology_Common_Versus_Unique_Elements_Across_Domains). Like, the aversion is due to focusing on different elements of the task.

>>>>

  Understanding the sporadic attractiveness of effortful options remains elusive. In this study, we tested the hypothesis that such appeal could arise from being able to compare different means to achieve the same outcome. Using two foraging experiments that exposed individuals to a single or intermixed types of demands, we found that 1) preferences for cognitive effort were stable when individuals face a single form of demand, but fade over time when other costs are also experienced; and 2) the phenomenon can be explained by a model in which preferences depend on how participants construe the amount of time a task takes, which could hypothetically be recalibrated when all tasks are interleaved.

  Rather than comparing discounting functions among demands, allowing people to choose for them helped us control for a number of confounds. For example, effortful tasks often take longer to perform in everyday life. We eliminated this possibility by using a common handling time for all demands.

  The observed choice dynamics ruled out a number of explanatory candidates. For instance, differences in choice stability between experiments suggest that a demand's cost is not intrinsic, while differences between effort types rule out learned industriousness. Effort preferences due to boredom would not have shown the global acceptance decline and convergence in the within-subjects experiment, which also discards the possibility that all our participants were high on need for cognition (CITE). Even though fatigue could explain the decrease in effort acceptances across experiments, we would have expected tired participants to turn to wait trials over time in order to preserve earning rates. We also ruled out the sunk cost effects used to explain overharvesting (Wikenheiser2013; Sweiss2018), as participants expressed their choices as early as possible, resulting in occassional underharvesting. Even though the reciprocal relationship between effort and outcome can often explain preferences for effort (Inzlicht2018), the observed differences in acceptances between effort types and the eventual convergence of rates in the within-subjects study cannot be explained by reward alone, as demands were rewarded at the same rate. Instead, we propose that features of the cognitive tasks added outcome-independent value to the offer. Subjective time sensitivity has been shown to play a role in temporal discounting, improving on classic hyperbolic fits (Mckelachar....) *add that stability meant no need to model asymmetric learning across environmental richness from Neal* (unlikely to be need for cognition bc of exp 2 global decrease)

# Points of discussion 
- We still think that cognitive and physical manipulations were effortful. Our physical effort task has been repeatedly used to probe effort. Similar task switching has been shown to feel cognitively effortful (Kool). Furthermore, the effortful nature of our manipulations is evidenced by the steeper decrease in acceptances over time at longer handling times (Treadway). There are many reasons why such cognitive manipulations can be effortful, which are often centered in the amount of control. Or the OC. 
- Note that the calibration of efforts were different. We adjusted to people's physical capabilities, but trained them in the cognitive tasks until they reached criterion. 
- Efficacy, along with reward amount, can influence how much control is exherted.
- But here we do not focus on choosing the *amount* of cognitive control to be exherted during performance, which can be due to efficacy, agency, or reward amount. Instead, we focus on what makes people *pursue* the task in spite of how effortful it is.
- Clinical: such deviations in the willingness to pursue effort can predict depression relapse (https://jamanetwork.com/journals/jamapsychiatry/article-abstract/2761562), and motivational factors affecting effort aversion can be heterogenous between depression and schizophrenia (https://www.researchgate.net/publication/337361164_Pathways_to_Motivational_Impairments_in_Psychopathology_Common_Versus_Unique_Elements_Across_Domains). Like, the aversion is due to focusing on different elements of the task.
- Having a between-subjects experiment provided a baseline of foraging behavior that helped us interpret the findings of the second experiment
- Control: for time as confound of effort
- Control: having two efforts helped dissociate domain-general from task-specific features.
- Control: participants were fully aware. This removed the possibility of asymmetric learning and choice difficulty.
- Control: in demand experiments, chosen options are often carried out after the experiment, but there are known discrepancies between expected and obtained value that could affect repeated decisions in real life.

- Control: even in TAFC tasks, selecting the shorter option could index
- Optimality let us know that people acted as if effort was attractive. This was evidenced by both the lower earnings by the cognitive effort group, and by the base OC model. Of note is that the optimality of the strategies produced by each cost depended on the richness of the block. This can help explain why the wait group earned the most despite having similar acceptance rates to the physical group, as the former was closer to optimality in the richest environment.
- This appeal couldn't be explained by a number of explanatory candidates
- This includes the often-cited reciprocal relationship between effort and reward
- But it turns out subjective time fits the bill. 
- Cognitive effort gained value by making trials feel shorter
- This could be because: 1) discrete nature of cog taska llowed for counting; 2) flow; 3) persistent recruitment of working memory (Buhi and Gershman). Talk about future experiments: if this is subjective time, changing the pacing of the tasks should change the acceptance rates for cognitive effort.
- maybe: gifts are valuable more when it required effort to acquire them.

  The observed choice dynamics ruled out a number of explanatory candidates. For instance, differences in choice stability between experiments suggest that a demand's cost is not intrinsic. Effort preferences due to boredom would not have shown the global acceptance decline and convergence in the within-subjects experiment, which also discards the possibility that all our participants were high on need for cognition (CITE). Even though fatigue could explain the decrease in effort acceptances across experiments, we would have expected tired participants to turn to wait trials over time in order to preserve earning rates. We also ruled out the sunk cost effects used to explain overharvesting (Wikenheiser2013; Sweiss2018), as participants expressed their choices as early as possible, resulting in occassional underharvesting. Even though the reciprocal relationship between effort and outcome can often explain preferences for effort (learned industriousness, Inzlicht2018), the observed differences in acceptances between effort types and the eventual convergence of rates in the within-subjects study cannot be explained by reward alone, as demands were rewarded at the same rate. Instead, we proposed that subjective time sensitivity has been shown to play a role in temporal discounting, improving on classic hyperbolic fits (Mckelachar....) *add that stability meant no need to model asymmetric learning across environmental richness from Neal* (unlikely to be need for cognition bc of exp 2 global decrease)

- Things that don't seem to be at play here, but could still play a general role in mental effort preferences: 1) A demand’s cost is not fully intrinsic (within vs between differences); 2) Effort preferences unlikely just due to learned industriousness or efficacy as amitais lab sees it (differences between effort types, and everyone met a good performance level per training); 3) avoiding boredom (global decline in exp 2, not a preference exchange). 4) Literally counting tasks, since mental effort was discretely presented but everything else was contiguous. If that were the case, dividing the handling time by 2 would work, but not sure if it would translated into exp 2. 5) fatigue: there's a similar gradual decrease in acceptances in both experiments, except for the wait condition in exp 1. Fatigue would apply equally to both, or alternatively it might not bring all down to the same level (I guess the point could be made, but it's not that people who complete more at the beginning decrease acceptances more, as one would expect from fatigued people). Also worth mentioning that even if this is due to need for cognition (i.e. Ceyda; Laura; Cacioppo et al 1996), the collapsing in the second experiment shows that preferences for cognitive effort cannot be just that.
